---
title: "ada master pipeline"
author: "Alec Plotkin"
date: '`r format(Sys.Date(), "%d %b %Y")`'
output: html_document
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this should cover my bases so that the working directory is correct when i run this markdown direct to console (as i usually do) or if i knit the entire thing
```

## libraries

- "readxl" is used for its `read_excel()` function, when extracting raw data
- "tidyverse" is used for dplyr's data wrangling functions, ggplot2's graphing functions, stringr's string manipulation functions, forcat's factor manipulation functions
- "knitr" is used for rmarkdown reports

```{r libs, warning = FALSE, message = FALSE}
# check if packages are installed and load libraries
if (!require(readxl)) install.packages('readxl')
library(readxl)
if (!require(knitr)) install.packages('knitr')
library(knitr)
if (!require(tidyverse)) install.packages('tidyverse')
library(tidyverse)
if (!require(here)) install.packages('here')
library(here)
```

## source functions
load functions to help processing

```{r functions}
source(here("ADA/R", "functions.R"))
source(here("ADA/R", "read_raw.R"))
```


## read in
extract plate maps and raw data from excel files.

```{r plate_maps, message = FALSE}
# identify directory with all files
dest <- 'ADA'

# read all plate maps in directory
map_files <- list.files(here(dest, "maps/plates"))
map_files <- map_files[!grepl("^~\\$", map_files)]  # delete temp files


# read in plate maps
maps <- purrr::map(here(dest, "maps/plates", map_files), 
                   read_excel_all_sheets, 
                   na.strings = c("", "NA", "---")) %>%
  flatten() %>%
  purrr::map(reformap)
```


## pipeline

the program lists datafiles in the `"raw"` folder, and uses their names to generate a list of `plate_names`.

```{r raw_files}
raw_files <- list.files(here(dest, "data/raw"))
raw_files <- raw_files[!grepl("^~\\$", raw_files)]  # delete temp files

# generate plate names by removing file extensions from data_files, clean up names, then order
raw_names <- sub(pattern = "(.*)\\..*$", replacement = "\\1", basename(raw_files))

if (length(maps) != length(raw_files))
  stop('The number of plates and plate maps is not equal')

# string to extract plate identifier
plate_id <- 'plate_\\d\\d'
raw_clean = stringr::str_extract(raw_names, plate_id)
maps_clean = stringr::str_extract(names(maps), plate_id)

# this needs to be turned into a tryCatch() statement...
if (!identical(raw_names, names(maps))) {
  warning('The names of raw files and plate maps do not match.\n\tRaw files:\n')
  show(raw_names)
  warning('\tPlate maps:\n')
  show(names(maps))
  if (identical(raw_clean, maps_clean)) {
    message('\nThe following names will be applied:')
    show(raw_clean)
  } else {
    stop('Matching names could not be found')
  }
}
```

extract absorbance change over time (delta) for all samples by sourcing the read_raw.R script in conjunction with sample names in plate maps

```{r read_excel, message = FALSE}
deltas <- purrr::pmap_dfr(
  list(platename = raw_names, platemap = maps),
  read_raw, 
  path = here(dest, "data/raw")
  )
```

read_raw.R saves each plate into its own R-readable datafile. we can read and combine these datafiles to get a tibble of all the samples' deltas

flag negative delta absorbances and set to NA:

```{r negatives}
# remove blanks from deltas, since they'll have separate qc:
blanks <- filter(deltas, grepl("Blk", sample_id))
deltas <- filter(deltas, !grepl("Blk", sample_id))

# create a column for flags w/in deltas
deltas$flag <- NA_character_

# check for negative values in averages
neg_abs <- deltas[which(deltas$dabs < 0), ]
cat("samples where average absorbance delta is below 0: ", nrow(neg_abs), "\\n")
neg_abs

deltas$flag[which(deltas$dabs < 0)] <- "dabs < 0"
deltas$dabs[which(deltas$dabs < 0)] <- NA_real_
message("averages below 0 set to NA")
```
now deltas can be used to prepare standard curves, following some QC.

```{r session_info}
sessionInfo()
```