---
title: "limits of detection"
author: "Alec Plotkin"
date: "1/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

calculate limit of detection for each plate using blank wells and the lowest standard(s), with the following equations:

```
# limit of blank:
lob = mean(blks) + 1.645*sd(blks)

# limit of detection:
lod = lob + 1.645*sd(lcs)
```

where lob = limit of blank, lod = limit of detection, and lcs = low concentration sample

blanks have already been extracted from data set of absorbance deltas; inspect them:

```{r blanks}
# look at overview of blanks
head(blanks, n = 20L)

# are there any negatives?
filter(blanks, dabs < 0)

# keep negatives (there are too many to exclude), but use them to calculate limit of blank:
lob <- blanks %>% 
  group_by(plate) %>% 
  summarize(lob = mean(dabs) + 1.645*sd(dabs))

lob  # no negatives here
```

need to extract low concentration samples... can find them in `qcdeltas` (since standards are subject to same pipetting errors as other samples, they are lumped into the same qc process). standard 7 is being used as the low concentration sample.

```{r lcs}
# do EHNA stds first:
lcs <- filter(qcdeltas, grepl("Std7$", sample_id))  # some samples had EHNA added to the standards... these samples are excluded
lcsd <- lcs %>%
  group_by(plate) %>%
  summarize(lcsd = sd(dabs, na.rm = TRUE))
```

combine blanks and lcs to calculate lod:

```{r lod}
lod <- merge(lob, lcsd) %>% 
  mutate(lod_dabs = lob + 1.645*lcsd,
         lod_labs = log(lod_dabs)) %>%
  select(-c("lob", "lcsd"))
lod
```

merge lod into qcdeltas and flag samples where labs < lod:

```{r lod_qc}
qcdeltas_lod <- merge(qcdeltas, lod, by = "plate")

index <- qcdeltas_lod %>%
  transmute(below = ifelse(labs < lod_labs, TRUE, FALSE)) %>%
  unlist() %>%
  as.logical() %>%
  which()
qcdeltas_lod$flag[index] <- "< LOD"

# store wells below lod in data frame
below_lod <- filter(qcdeltas_lod, flag == "< LOD")
# display total number below lod per plate
below_lod %>% group_by(plate) %>% summarize(n = n())
```

now samples can be fed into the standard curve

display session info

```{r session_info}
sessionInfo()
```